An AI Success Story in Metaphysics: \\ the Case of the Inconsistency in G\"odel's Ontological Argument

### 2016-01-11 Papers/InProgress/IJCAI/paper.tex
minor additions

Logic textbooks \cite{ToDo:which} commonly utilize classical higher-order logic (HOL)
\cite{Church} as a meta-language to introduce the syntax and the
semantics of object logics of interest, in which reasoning
problems in concrete application domains can be modeled and solved
with pen and paper. In fact, this approach can also be followed for even very challenging
object logics to
enable interactive and automated theorem proving with existing theorem provers for classical
higher-order logic.

@@

For a computational analysis of G\"odel's ontological argument in this
approach the embedding of higher-order modal logics (HOML) such as K,
KB and S5 with various domain conditions (possibilist and actualist)
is required. This idea has been successfully followed in related work
\cite{C40}. The embedding of higher-order modal logic is in fact
straightforward. Formulas in HMOL are \emph{lifted}, translated to predicates
over worlds, which are themselves explicitly represented as
terms. The logical constants of HMOL are translated to HOL terms in such a way that, for instance, 
%$\neg \varphi$, $\varphi\vee\psi$
$\Box \varphi$ and $\Diamond \varphi$ are mapped, respectively, to the HOL formulas $\forall w. (r w_0 w) \imp (\varphi w)$ and $\exists w. (r w_0 w) \wedge (\varphi w)$. This form of embedding is precisely the well-known standard translation \cite{Ohlbach,ModalLogicPatrickBlackburn},
which is here (intra-logically) realized in HOL by stating a set of
equations defining the logical constants (Fig. ~\ref{QML_S5}). The resulting object logic is the HOML K with rigid terms and constant domains (possibilist quantifiers). Other logics (e.g. KB, S5) can be embedded by adding axioms that restrict the accessibility relation $r$. Varying domains and actualist quantifiers can be simulated by means of an existence predicate, which can be used to guard the quantifiers. Therefore, the embedding approach is very flexible.

>>>>

Logic textbooks \cite{ToDo:which} commonly utilize classical higher-order logic (HOL)
\cite{Church40} as a meta-language to introduce the syntax and the
semantics of object logics of interest, in which reasoning
problems in concrete application domains can be modeled and solved
with pen and paper. In fact, this approach can also be followed for even very challenging
object logics to
enable interactive and automated theorem proving with existing theorem provers for classical
higher-order logic.

@@

For a computational analysis of G\"odel's ontological argument in this
approach the embedding of higher-order modal logics (HOML) such as K,
KB and S5 with various domain conditions (possibilist and actualist)
is required. This idea has been successfully followed in related work
\cite{C40}. The embedding of higher-order modal logic is in fact
straightforward. Formulas in HMOL are \emph{lifted}, translated to predicates
over worlds, which are themselves explicitly represented as
terms. The logical constants of HMOL are translated to HOL terms in such a way that, for instance, 
%$\neg \varphi$, $\varphi\vee\psi$
$\Box \varphi$ and $\Diamond \varphi$ are mapped, respectively, to the
HOL formulas $\forall w. (r w_0 w) \imp (\varphi w)$ and $\exists
w. (r w_0 w) \wedge (\varphi w)$. This form of embedding is precisely
the well-known standard translation
\cite{DBLP:journals/logcom/Ohlbach91}, which is here (intra-logically)
realized in HOL by stating a set of equations defining the logical
constants (Fig. ~\ref{QML_S5}). The resulting object logic is the HOML
K with rigid terms and constant domains (possibilist
quantifiers). Other logics (e.g. KB, S5) can be embedded by adding
axioms that restrict the accessibility relation $r$. Varying domains
and actualist quantifiers can be simulated by means of an existence
predicate, which can be used to guard the quantifiers. Therefore, the
embedding approach is very flexible.
